---
title: "A comparative analysis of regional economic resilience: employment rate and revised employment rate"
author: Jong-Seo Won
subtitle: Identifying resilient and non-resilient regions by using two different resilience indicators 
---

# Introduction
Recently, the notion of resilience has attracted attention from many economic geographers who try to understand how regional economies respond to and recover from disturbances. However, the concept of resilience in economic geography studies still remains unclear. Particularly, there is no universally-agreed resilience indicator to measure regional economic resilience.
In order to explore regional economic resilience of 51 regions in the U.S. in response to the globali financial crisis in the late 2000s, this study attempts to use different type of resilience indicator termed as 'revised employment rate', which is the ratio of employed population to total population from age 16 to 64 considered as working age, instead of employment rte, one of the most commonly used resilience indicators.


Employment rate = employed population / employed + unemployed population

Revised employment rate = employed population / age 15-64 population


The object of this project is to  explore regional resilience of 51 regions in the United States  by using two resilience indicators mentioned above. Then, visualize and compare outputs from two different resilience indicators; 'employment rate' and 'revised employment rate'



# Materials and methods
On the basis of engineering resilience perspective, it is determined whether a certain region is resilient or non-resilient in response to the economic crisis. If a region is resilient, it implies the region fully recovers from the impact of crisis. For example, a region's post-shock employment level bounces back and restores pre-shock level employment after the decline of employment due to the crisis. It is considered as a resilient region. Otherwise, a regional system does not restore its pre-shock state, it is considered as non-resilient region.



Narrative and most code will go here.  Describe what you are doing and show how to do it (with code).

You can do bullets like this:

* The first most important thing
* The second most important thing
* The third most important thing

You can do numbers like this:

1. The first most important thing
2. The second most important thing
3. The third most important thing

See [http://rmarkdown.rstudio.com/](http://rmarkdown.rstudio.com/) for all the amazing things you can do.


Here's my first code chunk.
```{r}


```

Load any required packages in a code chunk

```{r, message=F, warning=F}
library(readr)
library(sp)
library(acs)
library(censusapi)
library(tidycensus)


```




```{r, warning=F} 
## import dataset from Github

## import 2005 dataset (population, employment & unemployment)
ACS_05_EST_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_05_EST_S0701_with_ann.csv")

ACS_05_EST_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_05_EST_S2301_with_ann.csv")


## import 2006 dataset (population, employment & unemployment)
ACS_06_EST_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_06_EST_S0701_with_ann.csv")

ACS_06_EST_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_06_EST_S2301_with_ann.csv")


## import 2007 dataset (population, employment & unemployment)
ACS_07_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_07_1YR_S0701_with_ann.csv")

ACS_07_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_07_1YR_S2301_with_ann.csv")


## import 2008 dataset (population, employment & unemployment)
ACS_08_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_08_1YR_S0701_with_ann.csv")

ACS_08_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_08_1YR_S2301_with_ann.csv")


## import 2009 dataset (population, employment & unemployment)
ACS_09_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_09_1YR_S0701_with_ann.csv")

ACS_09_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_09_1YR_S2301_with_ann.csv")


## import 2010 dataset (population, employment & unemployment)
ACS_10_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_10_1YR_S0701_with_ann.csv")

ACS_10_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_10_1YR_S2301_with_ann.csv")


## import 2011 dataset (population, employment & unemployment)
ACS_11_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_11_1YR_S0701_with_ann.csv")

ACS_11_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_11_1YR_S2301_with_ann.csv")


## import 2012 dataset (population, employment & unemployment)
ACS_12_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_12_1YR_S0701_with_ann.csv")

ACS_12_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_12_1YR_S2301_with_ann.csv")


## import 2013 dataset (population, employment & unemployment)
ACS_13_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_13_1YR_S0701_with_ann.csv")

ACS_13_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_13_1YR_S2301_with_ann.csv")


## import 2014 dataset (population, employment & unemployment)
ACS_14_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_14_1YR_S0701_with_ann.csv")

ACS_14_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_14_1YR_S2301_with_ann.csv")


## import 2015 dataset (population, employment & unemployment)
ACS_15_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_15_1YR_S0701_with_ann.csv")

ACS_15_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_15_1YR_S2301_with_ann.csv")


## import 2016 dataset (population, employment & unemployment)
ACS_16_1YR_S0701_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_16_1YR_S0701_with_ann.csv")

ACS_16_1YR_S2301_with_ann <- read_csv("https://raw.githubusercontent.com/jongseow/RDataScience_Project/master/data_project/ACS_16_1YR_S2301_with_ann.csv")
```





## run the query and convert to data.frame()



# Results

Show tables, plots, etc. and describe them.

# Conclusions

What have you learned?  Are there any broader implications?

# References

